---
title: "Evaluation of performances"
output: 
  workflowr::wflow_html:
    includes:
      in_header: header.html
editor_options:
  chunk_output_type: console
author: "Patrick Schratz"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.retina = 3,
  fig.align = "center",
  fig.width = 6.93,
  fig.height = 6.13,
  out.width = "100%",
  echo = FALSE
)

R.utils::sourceDirectory("R")

# load drake objects
drake::loadd(
  bm_aggregated
)
library("mlr")
library("magrittr")
library("dplyr")
library("ggplot2")
library("ggrepel")
library("ggsci")
library("ggpubr")
```

## (Table) All leaner/filter/task combinations ordered by performance.

Taking into account all tasks, learners and filters: Who won?

```{r eval-performance-1, warning=FALSE}
getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
  mutate(task.id = recode_factor(task.id,
    `defoliation-all-plots-HR` = "HR",
    `defoliation-all-plots-VI` = "VI",
    `defoliation-all-plots-NRI` = "NRI",
    `defoliation-all-plots-HR-NRI` = "HR-NRI",
    `defoliation-all-plots-HR-VI` = "HR-VI",
    `defoliation-all-plots-HR-NRI-VI` = "HR-NRI-VI",
    
  )) %>%
  tidyr::separate(learner.id, c("learner_group", "filter"),
    remove = FALSE,
    sep = " "
  ) %>%
  group_by(learner_group, task.id, filter) %>%
  summarise(best = min(rmse.test.rmse)) %>%
  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
  arrange(best) %>%
  knitr::kable()
```

## (Table) Best learner/filter/task combination

Learners: On which task and using which filter did every learner score their best result on?

```{r eval-performance-2, warning=FALSE}
getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
  tidyr::separate(learner.id, c("learner_group", "filter"),
    remove = FALSE,
    sep = " "
  ) %>%
  group_by(learner_group) %>%
  slice(which.min(rmse.test.rmse)) %>%
  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
  arrange(rmse.test.rmse) %>%
  select(-6) %>%
  knitr::kable()
```

## (Plot) Best learner/filter combs for all tasks

```{r performance-results, warning=FALSE, dev = c("png", "pdf")}
results_aggr <- getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
  mutate(task.id = recode_factor(task.id,
    `defoliation-all-plots-HR` = "HR",
    `defoliation-all-plots-VI` = "VI",
    `defoliation-all-plots-NRI` = "NRI",
    `defoliation-all-plots-HR-NRI` = "HR-NRI",
    `defoliation-all-plots-HR-VI` = "HR-VI",
    `defoliation-all-plots-HR-NRI-VI` = "HR-NRI-VI",
  )) %>%
  tidyr::separate(learner.id, c("learner_group", "filter"),
    remove = FALSE,
    sep = " "
  ) %>%
  mutate(filter = replace(filter, is.na(filter), "NF")) %>%
  mutate(learner_group = recode_factor(learner_group, `XG` = "XGBOOST")) %>% 
  group_by(learner_group, task.id) %>%
  # get the best performance per learner and task
  # this is better than summarise() because we can keep additional columns
  # in constrast to summariuse which only keeps the grouping columns and the
  # summarised one
  slice(which.min(rmse.test.rmse))

results_aggr %>%
  ggplot(aes(x = rmse.test.rmse, y = task.id)) +
  geom_point(aes(color = learner_group), size = 2.5) +
  scale_color_nejm(breaks = sort(levels(results_aggr$learner_group))) +
  labs(x = "RMSE", y = "Task", color = "Learner") +
  guides(size = FALSE) +
  scale_x_continuous(limits = c(30, 70)) +
  # geom_label_repel(
  #   label = round(results_aggr$rmse.test.rmse, 3),
  #   size = 3, nudge_y = 0.3
  # ) +
  geom_label_repel(
    label = paste(results_aggr$filter,",",round(results_aggr$rmse.test.rmse, 3)),
    size = 2.8
  ) +
  theme_pubr()
```

## (Plot) Best filter combination of each learner vs. no filter per task

Showing the final effect of applying feature selection to a learner for each task.

```{r filter-effect, warning=FALSE, dev = c("png", "pdf")}
# we rbind two filtered data.frames
# doign the filtering in one step is complicated here


best_filter = getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
  mutate(task.id = recode_factor(task.id,
    `defoliation-all-plots-HR` = "HR",
    `defoliation-all-plots-VI` = "VI",
    `defoliation-all-plots-NRI` = "NRI",
    `defoliation-all-plots-HR-NRI` = "HR-NRI",
    `defoliation-all-plots-HR-VI` = "HR-VI",
    `defoliation-all-plots-HR-NRI-VI` = "HR-NRI-VI",
    
  )) %>%
  tidyr::separate(learner.id, c("learner_group", "filter"),
    remove = FALSE,
    sep = " "
  ) %>%
  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
  dplyr::filter(filter != "No Filter")  %>%
  group_by(learner_group, task.id) %>% 
  dplyr::filter(learner_group == "SVM" | learner_group == "RF") %>%
  slice(which.min(rmse.test.rmse))
    
best_no_filter = getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
  mutate(task.id = recode_factor(task.id,
    `defoliation-all-plots-HR` = "HR",
    `defoliation-all-plots-VI` = "VI",
    `defoliation-all-plots-NRI` = "NRI",
    `defoliation-all-plots-HR-NRI` = "HR-NRI",
    `defoliation-all-plots-HR-VI` = "HR-VI",
    `defoliation-all-plots-HR-NRI-VI` = "HR-NRI-VI",
    
  )) %>%
  tidyr::separate(learner.id, c("learner_group", "filter"),
    remove = FALSE,
    sep = " "
  ) %>%
  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
  dplyr::filter(filter == "No Filter")  %>%
  group_by(learner_group, task.id) %>% 
  dplyr::filter(learner_group == "SVM" | learner_group == "RF") %>%
  slice(which.min(rmse.test.rmse))
    
comb = bind_rows(best_filter, best_no_filter)
  
comb %>% 
  ggplot(aes(x = rmse.test.rmse, y = learner_group)) +
  geom_jitter(aes(color = filter), size = 2.5, width=0, height = 0.2) +
  facet_wrap(~ task.id) +
  scale_color_nejm() +
  labs(x = "RMSE", y = "Task", color = "Filter") +
  guides(size = FALSE) +
  scale_x_continuous(limits = c(30, 45)) +
  # geom_label_repel(
  #   label = round(results_aggr$rmse.test.rmse, 3),
  #   size = 3, nudge_y = 0.3
  # ) +
  # geom_label_repel(
  #   label = paste(comb$task.id),
  #   size = 2.8
  # ) +
  theme_pubr() +
  theme(panel.grid.major.y = element_line(size = 0.1, linetype = "dashed"))
```
