---
title: "Correlation of filter methods"
output: 
  workflowr::wflow_html:
    includes:
      in_header: header.html
editor_options:
  chunk_output_type: console
author: "Patrick Schratz"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.retina = 3,
  fig.align = "center",
  fig.width = 6.93,
  fig.height = 6.13,
  out.width = "100%",
  echo = FALSE
)

# source(here::here("code/99-packages.R"))
library("dplyr")
library("tidyr")
library("ggcorrplot")
library("purrr")
library("glue")
library("fs")
R.utils::sourceDirectory("R")

# load drake objects
drake::loadd(
  filter_values,
  filter_values_fselectorrcpp,

  filter_info_gain_nbins
)
```

# Correlation of filter methods {.tabset .tabset-fade}

Spearman's rank correlation is used because rankings are compared.

## VI

### Filter methods amongst each other

The idea behind is was to analyze the correlation between filter rankings.
We only wanted to included filters which have a somewhat unique ranking.
Otherwise, when creating ensemble filters, certain filters would implicitly be weighted more than others.

Takeaway:

- Only use one of "information gain", "gain ratio", "sym uncert"

- Either use Spearman or Pearson correlation

```{r filter-correlation-1, warning=FALSE}
fv_all_vi <- map(filter_values[8:14], ~ .x[["data"]]) %>%
  dplyr::bind_rows() %>%
  dplyr::select(value, filter) %>%
  mutate(id = rep(1:90, 7)) %>%
  tidyr::spread(filter, value) %>%
  dplyr::select(-id)

fv_all_vi_fselecrcpp <- map(filter_values_fselectorrcpp[4:6], ~ .x[["data"]]) %>%
  dplyr::bind_rows() %>%
  dplyr::select(value, filter) %>%
  mutate(id = rep(1:90, 3)) %>%
  tidyr::spread(filter, value) %>%
  dplyr::select(-id)

fv_all_vi_merged <- bind_cols(fv_all_vi, fv_all_vi_fselecrcpp)

corr_vi <- cor(fv_all_vi_merged, method = "spearman")

ggcorrplot(corr_vi,
  # p.mat = p.mat,
  hc.order = F,
  type = "lower",
  lab = TRUE
)
```

## NRI

```{r correlation-filter-nri, dev = c("png", "pdf"), warning=FALSE}
fv_all_nri <- map(filter_values[15:21], ~ .x[["data"]]) %>%
  dplyr::bind_rows() %>%
  dplyr::select(value, filter) %>%
  mutate(id = rep(1:7875, 7)) %>%
  tidyr::spread(filter, value) %>%
  dplyr::select(-id)

fv_all_nri_fselecrcpp <- map(filter_values_fselectorrcpp[7:9], ~ .x[["data"]]) %>%
  dplyr::bind_rows() %>%
  dplyr::select(value, filter) %>%
  mutate(id = rep(1:7875, 3)) %>%
  tidyr::spread(filter, value) %>%
  dplyr::select(-id)

fv_all_nri_merged <- bind_cols(fv_all_nri, fv_all_nri_fselecrcpp)

corr_nri <- cor(fv_all_nri_merged, method = "spearman")

ggcorrplot(corr_nri,
  # p.mat = p.mat,
  hc.order = F,
  type = "lower",
  lab = TRUE
)
```

## HR

```{r filter-correlation-4, warning=FALSE}
fv_all_hr <- map(filter_values[1:7], ~ .x[["data"]]) %>%
  dplyr::bind_rows() %>%
  dplyr::select(value, filter) %>%
  mutate(id = rep(1:122, 7)) %>%
  tidyr::spread(filter, value) %>%
  dplyr::select(-id)

fv_all_hr_fselecrcpp <- map(filter_values_fselectorrcpp[1:3], ~ .x[["data"]]) %>%
  dplyr::bind_rows() %>%
  dplyr::select(value, filter) %>%
  mutate(id = rep(1:122, 3)) %>%
  tidyr::spread(filter, value) %>%
  dplyr::select(-id)

fv_all_hr_merged <- bind_cols(fv_all_hr, fv_all_hr_fselecrcpp)

corr_hr <- cor(fv_all_hr_merged, method = "spearman")

ggcorrplot(corr_hr,
  # p.mat = p.mat,
  hc.order = F,
  type = "lower",
  lab = TRUE
)
```

## Number of bins of `FSelectorRcpp::information.gain()`

Analyzing the effect of a different `nbins` value on the filter values of filter "Information Gain".

- Lower correlation / highest difference: `nbins = 5` vs. `nbins = 30`

-> We decided to use with `nbins = 10` in the analysis.

The hidden default of nbins when setting `equal = TRUE` in `FSelectorRcpp::information_gain()` is 5.

```{r correlation-nbins, warning=FALSE, dev = c("png", "pdf")}
# recode levels so that they are unique
fv_info_gain <- map(filter_info_gain_nbins[1:6], ~ .x[["data"]]) %>%
  dplyr::bind_rows() %>%
  dplyr::select(value, filter) %>%
  mutate(id = rep(seq(5, 30, 5), each = 122)) %>%
  tidyr::unite(filter, c("filter", "id")) %>%
  mutate(id = rep(seq(1, length(.data$value) / 6), 6)) %>%
  tidyr::pivot_wider(
    names_from = filter,
    values_from = value
  ) %>%
  dplyr::select(-id)

corr_inf_gain <- cor(fv_info_gain, method = "spearman")

ggcorrplot(corr_inf_gain,
  # p.mat = p.mat,
  hc.order = F,
  type = "lower",
  lab = TRUE
)
```

