---
title: "Checking for normality of response variable"
output: 
  workflowr::wflow_html:
    includes:
      in_header: header.html
editor_options:
  chunk_output_type: console
author: "Patrick Schratz"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.retina = 3,
  fig.align = "center",
  fig.width = 6.93,
  fig.height = 6.13,
  out.width = "100%",
  echo = FALSE
)
library("drake")
library("ggplot2")
library("MASS")

loadd(
  vi_data_corrected_buffer2,
  benchmark_models_new_penalized_buffer2,
  benchmark_models_new_buffer2
)
```

Last update: 

```{r}
date()
```

This document originated from the fear of having a response variable which is not normally distributed "enough".

The response variable looks as follows:

```{r response-normality-1}
ggplot(vi_data_corrected_buffer2, aes(x = defoliation)) +
  geom_histogram(bins = 30)
```

When applying the [Shapiro-Wilk](https://en.wikipedia.org/wiki/Shapiro%E2%80%93Wilk_test) test we get

```{r response-normality-2}
shapiro.test(vi_data_corrected_buffer2$defoliation)
```

## Exploring model residuals 

Visualizing model residuals of LASSO and RF to see how they differ.
The LASSO "predicted vs. fitted" plot shows limited model power.

### Lasso model with no transformation of the response variable

```{r response-normality-3}
pred <- benchmark_models_new_penalized_buffer2[[1]]$results$hr$`Lasso-CV`$pred

# fit lm of y and 天
resid <- lm(pred$data$truth ~ pred$data$response)

# https://stats.stackexchange.com/a/53257/101464
# plot(density(resid(resid)))

par(mfrow = c(1, 2))
plot(resid, which = c(1:2), ask = FALSE, main = "No Transform")
par(mfrow = c(1, 1))


# without outlier
resid1 <- resid
resid1$fitted.values[620] <- NA

par(mfrow = c(1, 2))
plot(resid1, which = c(1:2), ask = FALSE, main = "No Transform")
par(mfrow = c(1, 1))
```

### SVM with no transformation of the response variable

```{r response-normality-4}
pred_svm <- benchmark_models_new_buffer2[[1]]$results$hr$`SVM MBO MRMR`$pred

# fit lm of y and 天
resid_svm <- lm(pred_svm$data$truth ~ pred_svm$data$response)

# https://stats.stackexchange.com/a/53257/101464
# plot(density(resid(resid)))

par(mfrow = c(1, 2))
plot(resid_svm, which = c(1:2), ask = FALSE, main = "No Transform")
par(mfrow = c(1, 1))
```

# Variable Transformations

The following transformations of the response variable were done to check if it they have an effect on the "residuals vs. fitted" and "QQ-Plot" shown above.

## Power transformation

One option to enforce more normality of a variable is by applying a power transformation.
The [Box-Cox](https://en.wikipedia.org/wiki/Power_transform#Box%E2%80%93Cox_transformation) power transformation estimates a lambda value from the variable.
Next, the transformation can be applied via

$$(y^lambda - 1) / lambda$$

There is a [Stackoverflow question](https://stackoverflow.com/questions/33999512/how-to-use-the-box-cox-power-transformation-in-r) that shows how to do this. 

### Box-Cox Transformation

Applying it on the response of the data.

```{r response-normality-5}
n <- 1808
x <- runif(n, 1, 5)
y <- vi_data_corrected_buffer2$defoliation

# 1st approach to estimate lambda for power transformation
bc <- boxcox(y ~ x)
(lambda <- bc$x[which.max(bc$y)])

# apply power transform (boxcox)
y_new1 <- (y^lambda - 1) / lambda
shapiro.test(y_new1)
```

The "W" value is a little less than before.

There is another way to do this via package _car_.

```{r response-normality-6}
# 2nd approach to estimate lambda for power transformation

# boxcox (0.84)
y_new2 <- (y^(car::powerTransform(y)$lambda) - 1) / car::powerTransform(y)$lambda
shapiro.test(y_new2)
```

Exploring the residuals of a Lasso model with no transformation of the response variable.

```{r response-normality-7}
pred <- benchmark_models_new_penalized_buffer2[[1]]$results$hr$`Lasso-CV`$pred

# fit lm of y and 天
resid <- lm(pred$data$truth ~ pred$data$response)

# https://stats.stackexchange.com/a/53257/101464
# plot(density(resid(resid)))

# residuals vs. fitted & QQ plot
par(mfrow = c(1, 2))
plot(resid, which = c(1:2), ask = FALSE, main = "No Transform")
par(mfrow = c(1, 1))
```

<!-- ### Tukey Transformation -->

<!-- A slightly different way is to se the so called "Tukey" transformation instead of the "Box-Cox" transformation. -->

<!-- ```{r response-normality-8} -->
<!-- # tukey trans (0.846) -->
<!-- y_new3 <- y^(car::powerTransform(y)$lambda) -->
<!-- shapiro.test(y_new3) -->
<!-- ``` -->

<!-- ## Log transform -->

<!-- Beforehand we did a log transformation of the data. -->
<!-- However, since the data is a bit lef-skewed, this was enforced even more by that operation. -->
<!-- A substantially lower "W" value is the result. -->

<!-- ```{r response-normality-9} -->
<!-- # log transformed response (0.53) -->
<!-- shapiro.test(log(y)) -->
<!-- ``` -->

<!-- Also if we take a look at the residuals, they do not show normality. -->

<!-- ```{r response-normality-10} -->
<!-- pred_bc <- bm_models_vi_task_lasso_no_filter_mbo$results$`defoliation-all-plots-VI`$lasso$pred -->

<!-- # fit lm of y and 天 -->
<!-- resid_bc <- lm(pred_bc$data$truth ~ pred$data$response) -->

<!-- # https://stats.stackexchange.com/a/53257/101464 -->
<!-- # plot(density(resid(resid_bc))) -->

<!-- par(mfrow = c(1, 2)) -->
<!-- plot(resid_bc, which = c(1:2), ask = FALSE, main = "Boxcox") -->
<!-- par(mfrow = c(1, 1)) -->
<!-- ``` -->

<!-- Also the Shapiro test does not look good -->

<!-- ```{r response-normality-11} -->
<!-- shapiro.test(resid$residuals) -->
<!-- ``` -->

<!-- ## Inverse Box-Cox Transformation -->

<!-- Just for completeness. -->

<!-- ```{r response-normality-12} -->
<!-- y_old <- (lambda * y_new1 + 1)^(1 / lambda) -->
<!-- ``` -->

